---
title: "Examples for Third-Party Packages"
author: "Lukasz A. Bartnik"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples for Third-Party Packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Deferred and Remote

In the [Introduction](intro.html) vignette we explain the steps to create
a simple __deferred function wrapper__. Here we will show how it can be
combined with other packages to implement a simple RPC mechanism.


## Preparing the Wrapper

Unlike in the [Introduction](intro.html), to avoid clouding the message,
this time we will work with a much simplified example. Here's the function
we want to run in a remote R session:

```{r original}
fun <- function(x, y) x ^ y
```

Now we need to wrap it using `defer()`. Rrunning the wrapper is still
straightforward.

```{r}
library(defer)

deferred_fun <- defer(fun)
deferred_fun(9, 2)
```

Because it is easier to run an argument-less function, our first step
will be to __augment__ our deferred function wrapper with a value for
both of its arguments.

```{r}
deferred_fun <- augment(deferred_fun, x = 7, y = 3)
```

Now we can call `deferred_fun()` completely ommitting arguments (see
`?defer::augment` for more details).

```{r}
deferred_fun()
```



## OpenCPU

Our first example will be [OpenCPU](www.opencpu.org). We will also need
two other packages: [jsonlite](https://cran.r-project.org/package=jsonlite)
to serialize the wrapper and [httr](https://cran.r-project.org/package=httr)
to talk to OpenCPU's HTTP-based API. `jsonlite` is a prerequisite of
`opencpu` so we can be sure it will be available.


### Making the Call

With `OpenCPU` we can run any R function available in any of the packages
installed on the remote R server. `OpenCPU` also knows how to handle an
uploaded file, thus we will call `base::source()` and send a very simple
script, `local_script_path`, as its argument. We will call a public
`OpenCPU` server, __cloud.opencpu.org__.

```{r}
library(httr)
library(jsonlite)

public_opencpu_url <- "https://cloud.opencpu.org/ocpu/library/base/R/source/print"
local_script_path <- tempfile(fileext = ".R")
```


Here is how we invoke the remote `OpenCPU` server throgh `httr`:

```{r opencpu_call,eval=FALSE}
http_result <- httr::POST(public_opencpu_url,
                          body = list(file = upload_file(local_script_path)))
```


### The Script

The script contains `deferred_fun` serialized into a base64-encoded string.
It instructs the remote R session to decode the string, unserialize the
wrapper and then run it in that remote R session.

```{r opencpu_script}
serialized_wrapper <- jsonlite::base64_enc(serialize(deferred_fun, NULL))
cat(paste0("bytes <- unserialize(jsonlite::base64_dec('", serialized_wrapper, "'))\n",
           "deserialized()\n"),
    file = local_script_path)
```

### Running the OpenCPU Example

Now that everything is set, we can run the HTTP call and collect
the result.

```{r}
http_result <- httr::POST(public_opencpu_url,
                          body = list(file = upload_file(local_script_path)))
content(http_result, 'text')
```

`source()` returns a `list` with two elements: `value` is the value of
the last statement in the file - which is 7 to the power of 3, as
expected; `visible` is an attribute that tells us if `value` was intended
to be printed out.



## foreach

## Rserve

## SparkR

## (planned) extension for dplyr::do
